# Bias  {-}

---

```{r, fig.cap="[@Baker_2023-cy](https://stanforddaily.com/2023/02/17/internal-review-found-falsified-data-in-stanford-presidents-alzheimers-research-colleagues-allege/)"}
knitr::include_url("https://stanforddaily.com/2023/02/17/internal-review-found-falsified-data-in-stanford-presidents-alzheimers-research-colleagues-allege/", height = "600px")
```

## Roadmap {-}

- Discuss
    - [@Ritchie2020-fm], Chapter 4
- Discuss results from
    - [Exercise 02: P-hack your way to scientific glory](https://psu-psychology.github.io/psych-490-reproducibility-2023-spring/ex_p_hacking).
- [Due today]{.orange_due}
    - [Exercise 02: P-hack your way to scientific glory](https://psu-psychology.github.io/psych-490-reproducibility-2023-spring/ex_p_hacking)

## Kinds of bias {-}

- Analysts' prior assumptions
- Researcher, reviewer, and publication bias toward positive (vs. negative/null) results
    - File-drawer effect
- Preference for simple results (c.f., Stapel)
- Overfitting
- Conflicts of interest
- *p*-hacking

## Discuss p-hacking exercise {-}

1. Who got a "significant" result?
1. How many different analyses did you try?
1. Who changed their analysis after finding a significant result?
1. Did anyone try another analysis--after you got a significant result--and *keep* the non-significant result?

### Preparation {-}

It often saves typing to load a set of commands into memory. In R, groups of useful commands are called 'packages'. We can load a set of useful packages into memory by issuing the following command.

```{r}
library(tidyverse)
```

If you are interested in a career related to data science, [`tidyverse`](https://www.tidyverse.org) is a very powerful set of tools you will want to know more about.

### Data entry {-}

Via a Google Sheet: <https://docs.google.com/spreadsheets/d/1fnSwFrUcKvgqq_agDLe4t2DHXHtHoOlmLdtLVRSemrI/edit?usp=sharing>

::: {.rmdnote}

Gilmore added data validation (Format/Data Validation) to the columns. Why?

:::

::: {.rmdnote}

These data are "long".
Each row is a unique observation.
Long data are often easier to work with.
But not always.

:::

### Data gathering {-}

First, I authenticate (sign-in) to Google using my Gmail account. If I haven't logged in using this script recently, it will ask me to log-in again.

```{r}
googledrive::drive_auth("rick.o.gilmore@gmail.com")
```

Then I download the Google Sheet to a directory/folder called `csv/` using the file name `p-hacking.csv`.


```{r}
googledrive::drive_download(file = "PSYCH 490.002 2023 P-hacking", path = "csv/p-hacking.csv", type = 'csv', overwrite = TRUE)
```

::: {.rmdnote}

What does CSV mean?

Why are CSV files often used in data analysis?

:::

```{r}
p_hacking <- read_csv(file = "csv/p-hacking.csv", show_col_types = FALSE)
```

Next, I read the CSV file using the `read_csv()` function.

Functions in R take inputs and deliver outputs. The inputs are `file` and `show_col_types`.

The output is an object called `p_hacking`. It is a table of data that I can refer to with that name.

I like to use the 'structure' function or `str()` to see what the data look like.

::: {.rmdnote}

Data is a plural noun. So, (when we don't forget this) we say 'The data are...' not 'The data is...'.

:::

```{r}
str(p_hacking)
```

### Questions to explore {-}

Most data analysts find that the process of exploring data is iterative.

We start with a question. That leads to another question. That leads to yet another question.

It is also sometimes cyclical. To answer a question requires that we modify the form of our data file.

I like to start with thinking about "data pictures." If X was true, what would the data look like?

---

The following code we may or may not use. I put it here so it's easier for all of us if we need to make use of it.

```{r make-power-long, eval=FALSE}
power_df <- p_hacking |>
  pivot_longer(cols = contains('power_'), 
                      names_to = "political_positions", 
                      values_to = "pol_pos_selected") |>
  distinct() |>
  mutate(political_positions = stringr::str_remove(string = political_positions,
                                                             pattern = "power_"))
  
```

```{r make-econ-long, eval=FALSE}
econ_df <- p_hacking |>
  pivot_longer(cols = contains('econ_'), 
                      names_to = "econ_measures", 
                      values_to = "econ_meas_selected") |>
  distinct() |>
  mutate(econ_measures = stringr::str_remove(string = econ_measures,
                                                             pattern = "econ_"))
  
```

### Visualize {-}

```{r hist-p-vals}
p_hacking |>
  ggplot() +
  aes(x = p_value, fill = party) +
  geom_histogram() +
  facet_grid(~ prediction)
```

```{r p-vals-by-analysis}
p_hacking |>
  ggplot() +
  aes(x = analysis, y = p_value) +
  geom_point() +
  geom_line()
```

## Next time... {-}

- File drawer effect
    - [@Rosenthal1979-zi](https://doi.org/10.1037/0033-2909.86.3.638)
    - [@Franco2014-yu](http://dx.doi.org/10.1126/science.1255484)
- Work session on final project [proposals](https://psu-psychology.github.io/psych-490-reproducibility-2023-spring/final_project.html), [due Thursday, March 2]{.orange_due}.
